---
title: "tutorial 8"
author: "jared sleppy"
date: "10/25/2019"
output: html_document
---




## set up

In this dataset, I will be analyzing the traffic stops in Wichita, KS.

```{r load-data}
library(tidyverse)
stops <- read_csv("https://datajournalism.tech/wp-content/uploads/2019/10/wichita.csv")

population_2016 <- tibble(subject_race = c("asian/pacific islander", "black", "hispanic", "other/unknown", "white"), num_people = c(19294, 42485, 65090, 16686, 245499)) %>% mutate(subject_race = as.factor(subject_race))

center_lat <- 37.685260
center_lng <- -97.336411
```

## exploratory data analysis

You can also embed plots, for example:

```{r analysis, echo=FALSE}
colnames(stops)
nrow(stops)
str(stops)
summary(stops)
race_group <- stops %>% 
  group_by(subject_race) %>% 
  summarize(n = n(), prop = n / nrow(.))
race_group
```

To conduct a benchmark test, I will compute the proportions of demographic data in Wichita, KS.
```{r}
population_2016 <- population_2016 %>% 
  mutate(prop = num_people / sum(num_people))
```


Next step I will combine the two tables of population_2016 and race_group to conduct a benchmark test.
```{r message=F}
stops_final <- race_group %>% 
  left_join(population_2016, by = "subject_race") %>% 
  mutate(stop_rate = n / num_people)
  
```

## Data Visualization
I will visualize that stop rate by the police in Wichita, KS using a bar chart.

```{r}
bar <- ggplot(stops_final, aes(x=reorder(subject_race,stop_rate), y = stop_rate))+
  geom_bar(stat="identity", position="identity", fill="pink")+
  labs(title="Stopped Drivers by Race", subtitle = "African American drivers got stopped more than White drivers in the city of Wichita, Kansas")+
  coord_flip()+
geom_label(aes(x= subject_race, y=stop_rate, label= round(stop_rate,2)))
bar
```

Here is an interactive map. click on the dots to see more details about the drivers.
```{r}
library(httpuv)
library(leaflet)
race <- colorFactor(c("pink", "blue", "white", "burlywood3", "green"),
          domain = c("asian/pacific islander", "black", "white", "hispanic", "other/unknown"), ordered = TRUE)
f <- stops %>% drop_na(lat, lng)
map <- leaflet(f) %>% 
  addProviderTiles(providers$CartoDB) %>% 
  setView(lng = center_lng, lat = center_lat, zoom = 16) %>% 
  addCircleMarkers(~lng,
                   ~lat,
                   popup=paste("Race:",  f$subject_race, "Sex:", f$subject_sex, "Age:", f$subject_age),
                   weight = 3,
                   radius = 4,
                   color=~race(subject_race),
                   stroke=F,
                   fillOpacity =1)
map
```


